### 2021-06-07 thesis/tex/sterileResults.tex
update

% \begin{document}
\section{The Sterile Hypothesis}
With the null ($3\nu$ hypothesis) normalized to the IceCube Monte Carlo as in Eq.~\ref{eq:Nth}, we are now in good shape to study the sterile effect on the probabilities, and how that compares to data.
First, let us see how the collected data from Fig.~\ref{fig:IC_data} deviates from the predicted $3\nu$ oscillations.
\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{figures/IC_rates.pdf}
    \caption{}\label{fig:IC_rates}
\end{figure}
Fig.~\ref{fig:IC_rates} shows an impressive agreement to the standard $3\nu$ oscillation picture, with the largest deviations of $~6\%$
for neutrinos close to double-digit \si{\TeV} values. In zenith, the data only deviates $\pm 1\%$. Recalling the \si{\TeV} disappearance from 
Fig.~\ref{fig:sterile_resonance}, we see that we have a similar deficiency found in the data at \SI{4}{\TeV}. Thus, we expect the best fitting 
sterile hypothesis to include a $\dm[41]$ that places the resonant disappearance in the same region. From Fig.~\ref{fig:resonance_shift}, we see
that a mass of $\dm[41] = \SI{1.5}{\eV \squared}$ achieves this.

\subsection{$\chi^2$ minimization}
For our analyses, we define our $\chi^2$ as
\begin{align} \label{eq:chisq}
    \chi^{2}(\hat{\theta},\alpha,\beta)=\sum_{ij} \frac{\left(N^\text{th}-N^\text{data}\right)_{ij}^{2}}
    {\left(\sigma^\text{data}_{ij}\right)^{2} + \left(\sigma^\text{syst}_{ij}\right)^{2}}+ 
    \frac{(1-\alpha)^2}{\sigma_\alpha^2} + \frac{\beta^2}{\sigma_\beta^2}\,
\end{align}
where we minimize over the model parameters $\hat{\theta}$, the penalty terms $\alpha$ and $\beta$.
$N_{ij}^\text{th}$ is the expected number of events from theory, and $N_{ij}^\text{data}$ is the observed number of events in that bin. 
We set $\sigma_\alpha = 0.25$ as the atmospheric flux normalization error, and $\sigma_\beta = 0.04$ as the zenith angle slope error~\cite{hondapaper}. 
The observed event number has an associated Poissonian uncertainty $\sigma_{ij}^\text{data} = \sqrt{N_{ij}^\text{data}}$.
For IceCube, the event count takes the form
\begin{align}
    N^\text{th}_{ij} = \alpha\left[1+\beta (0.5 + \zreco_i )\right] N_{ij}(\hat{\theta})\,,
\end{align}
with $N_{ij}(\hat{\theta})$ from Eq.~\ref{eq:Nth}. Here, the term $ \beta (0.5 + \zreco_i )$ allows the event distribution to rotate with angle $\beta$ around the median zenith angle of $\zreco = -0.5$.
We also have an uncorrelated systematic error $\sigma_{ijk}^\text{syst}$. We set $\sigma_{ijk}^\text{syst} = f\sqrt{N_{ijk}^\text{data}}$, where $f$ is a percentage of our own choosing,
typically a value between 0 and 20 \%.

>>>>

% \begin{document}
\section{The Sterile Hypothesis}
With the null ($3\nu$) hypothesis normalized to the IceCube Monte Carlo as in Eq.~\ref{eq:Nth}, we are now in good shape to study the sterile effect on the probabilities, and how that compares to data.
First, let us see how the collected data from Fig.~\ref{fig:IC_data} deviates from the predicted $3\nu$ oscillations.
\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{figures/IC_rates.pdf}
    \caption{}\label{fig:IC_rates}
\end{figure}
Fig.~\ref{fig:IC_rates} shows an impressive agreement to the standard $3\nu$ oscillation picture, with the largest deviations of $~6\%$
for neutrinos close to double-digit \si{\TeV} values. In zenith, the data only deviates $\pm 1\%$. Recalling the \si{\TeV} disappearance from 
Fig.~\ref{fig:sterile_resonance}, we see that we have a similar deficiency found in the data at \SI{4}{\TeV}. Thus, we expect the best fitting 
sterile hypothesis to include a $\dm[41]$ that places the resonant disappearance in the same region. From Fig.~\ref{fig:resonance_shift}, we see
that a mass of $\dm[41] = \SI{1.5}{\eV \squared}$ achieves this.

\subsection{\texorpdfstring{$\chi^2$}{Chi-squared} Minimization}
For our analyses, we define our $\chi^2$ as
\begin{align} \label{eq:chisq}
    \chi^{2}(\hat{\theta},\alpha,\beta)=\sum_{ij} \frac{\left(N^\text{th}-N^\text{data}\right)_{ij}^{2}}
    {\left(\sigma^\text{data}_{ij}\right)^{2} + \left(\sigma^\text{syst}_{ij}\right)^{2}}+ 
    \frac{(1-\alpha)^2}{\sigma_\alpha^2} + \frac{\beta^2}{\sigma_\beta^2}\,
\end{align}
where we minimize over the model parameters $\hat{\theta}$, the penalty terms $\alpha$ and $\beta$.
$N_{ij}^\text{th}$ is the expected number of events from theory, and $N_{ij}^\text{data}$ is the observed number of events in that bin. 
We set $\sigma_\alpha = 0.25$ as the atmospheric flux normalization error, and $\sigma_\beta = 0.04$ as the zenith angle slope error~\cite{hondapaper}. 
The observed event number has an associated Poissonian uncertainty $\sigma_{ij}^\text{data} = \sqrt{N_{ij}^\text{data}}$.
For IceCube, the event count takes the form
\begin{align}
    N^\text{th}_{ij} = \alpha\left[1+\beta (0.5 + \zreco_i )\right] N_{ij}(\hat{\theta})\,,
\end{align}
with $N_{ij}(\hat{\theta})$ from Eq.~\ref{eq:Nth}. Here, the term $ \beta (0.5 + \zreco_i )$ allows the event distribution to rotate with angle $\beta$ around the median zenith angle of $\zreco = -0.5$.
We also have an uncorrelated systematic error $\sigma_{ijk}^\text{syst}$. We set $\sigma_{ijk}^\text{syst} = f\sqrt{N_{ijk}^\text{data}}$, where $f$ is a percentage of our own choosing,
typically a value between 0 and 20 \%.

### 2021-06-07 thesis/tex/sterileTheory.tex
update

% \input{commands}\begin{document}
\section{The Sterile State}\label{sec:anomalies}
In 1996, the LSND experiment reported an excess of $\ane$ events from an $\anm$ beam~\cite{lsnd}.  
Nine years later, MiniBooNE not only reproduced the $\ane$ anomaly, but saw the excess in The
$\ne$ events too. Together with the so-called reactor and gallium anomalies, these reports suggested 
that the three massive neutrino framework could be amended.
By introducing a fourth heavy mass state $\nu_4$, both appearance and disappeareance anomalies could be minimally accomodated.

However, we know from the decay width of the $Z$ boson that it only can interact with three flavor species, 
so this fourth mass state cannot be interacting weakly.
We now distinguish between the three original neutrino flavors ($e$, $\mu$, and $\tau$) and the new fourth 
flavor ($s$) by calling the former \emph{active} neutrinos
and the latter \emph{sterile}. The experiments listed above indicate that the mass-squared difference of the 
sterile neutrino is in the \si{\eV} scale, while the two others
are three and five magnitudes smaller. To remind us of this difference, we write $3+1$, since the sterile state 
proposed is a lot more massive than the active states. Models
with different hierachies might be written as $1+2+1$ or $1+3$, but many of them are ruled out by cosmological 
observations and will not be considered here. %TODO:source

In this chapter, we will use the matter-enhanced resonant oscillations assuming the 3+1 framework,
and probe it the using the $\nm$ disappearance signature.

@@

% \newpage\bibliographystyle{elsarticle-num}\bibliography{ref}\end{document}

>>>>

%\input{commands}
%\begin{document}
\section{The Sterile State}\label{sec:anomalies}
In 1996, the LSND experiment reported an excess of $\ane$ events from an $\anm$ beam~\cite{lsnd}.  
Nine years later, MiniBooNE not only reproduced the $\ane$ anomaly, but saw the excess in The
$\ne$ events too. Together with the so-called reactor and gallium anomalies, these reports suggested 
that the three massive neutrino framework could be amended.
By introducing a fourth heavy mass state $\nu_4$, both appearance and disappeareance anomalies could be minimally accomodated.

However, we know from the decay width of the $Z$ boson that it only can interact with three flavor species\footnote{
    Direct measurements of the invisible $Z$ decay yields $N=2.92 \pm 0.05$\cite{pdg}.
}, 
so this fourth mass state cannot be interacting weakly. In other words, it needs to transform as a
singlet under the broken gauge group $SU(2)_L \times U(1)_{EM}$.
We now distinguish between the three original neutrino flavors ($e$, $\mu$, and $\tau$) and the new fourth 
flavor by calling the former \emph{active} neutrinos
and the latter \emph{sterile} due to its non-interacting behavior.

The experiments listed above indicate that the mass-squared difference of the 
sterile neutrino is in the \si{\eV\squared} scale, while the two others
are three and five magnitudes smaller. To remind us of this difference, we write $3+1$, since the sterile state 
proposed is a lot more massive than the active states.

Two neutrino physics parameters probed by CMB lensing observations are the effective number of neutrinos
$N_\text{eff}$, and the sum of neutrino masses $\sum m_i$. 
The 2018 results from the Planck Collaboration in~\cite{planck2018} constrain the parameters to 
\begin{align}
    N_\text{eff} = 2.96^{+0.34}_{-0.33}\,, \quad \sum m_i < \SI{0.12}{\eV} \quad (95\%)\,,
\end{align}
consistent with the Standard Model with oscillations\footnote{Of course, the 
Standard Model \emph{without} neutrino masses dictates $N_\text{eff} = 3$ and $\sum m_i \equiv 0$.} prediction of $N_\text{eff} = 3.045$ degrees~\cite{desalas2016}.
Assuming the Planck data, one thermalized sterile neutrino $(N_\text{eff} = 4)$ is excluded at $6\sigma$. 
Thus, thermalized sterile 
neutrinos are strongly in tension with cosmological observations if one does not consider some special mechanism.
So if a fourth neutrino species is present in nature, it will have ramifications not only on neutrino physics itself, but on
our understanding of the CMB aswell.

@@

%\newpage\bibliographystyle{elsarticle-num}\bibliography{ref}\end{document}
